2.6: Inference for Regression

### 2019-10-08 static/slides/12-slides.Rmd
Tweak slides 12
Tweaks to slides 12 (formatting fixes, and some more about p-values)

- We want to .whisper[identify] causal relationships between **population** variables
    - Logically first thing to consider
    - **Endogeneity problem**

- We'll use **sample** *statistics* to .whisper[infer] something about population *parameters*
    - In practice, we'll only ever have a finite *sample distribution* of data
    - We *don't* know the *population distribution* of data
    - **Randomness problem**

@@

- .whisper[All modern science is built upon statistical hypothesis testing, so understand it well!]

@@

- Modern philosophy of science is largely based off of hypothesis testing and .onfire[falsifiability], which form the "Scientific Method"<sup>1</sup>

- For something to be "scientific", it must be .whisper[falsifiable], or at least .onfire[testable]

@@

- .whisper[Test statistic `\\((\delta)\\)`]: measures **how far what we observed in our sample $(\hat{\beta_1})$ is from what we would expect if the null hypothesis were true $(\beta_1=0)$**
    - Calculated from a sampling distribution of the estimator (i.e. $\hat{\beta_1})$
    - In econometrics, we use $t$-distributions which have $n-k-1$ degrees of freedom<sup>.red</sup>

- .whisper[Rejection region]: if the test statistic reaches a .onfire["critical value"] of $\delta$, then we **reject** the null hypothesis

@@

```{r}
# save as obs_slope
sample_slope <- school_reg_tidy %>% # this is the regression tidied with broom
  filter(term=="str") %>%
  pull(estimate)

@@

```{r, echo=T, eval=F}
CASchool %>%
  specify(testscr ~ str)%>%
  head(.,3)
```

@@

```{r}
CASchool %>%
  specify(testscr ~ str) %>%
  hypothesize(null = "independence") %>%
  head(.,3)
```
]

.footnote[<sup>.red[1]</sup> `type` can be either `point` (for specific point estimates for a single variable, such as a sample mean, `\\((\bar{x})\\)`, or `independence` (for hypotheses about two samples or a relationship between variables). See more [here](https://moderndive.netlify.com/9-hypothesis-testing.html).]
---

@@

- onfire[Two-sided alternative] $H_a: \beta_1 \neq 0$, we double the raw $p$-value

@@

.footnote[<sup>1</sup> $k$ is the number of $X$ variables.]

@@

```{r, fig.retina=3, fig.height=5, fig.align="center"}
ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  stat_function(fun=dnorm, xlim=c(-4,-2), geom="area", fill="blue", alpha=0.5)+
  stat_function(fun=dnorm, xlim=c(2,4), geom="area", fill="blue", alpha=0.5)+
  labs(x = "t",
       y = "Probability")+
  scale_x_continuous(breaks=c(-2,2),
                      label=c(expression(-t[i]), expression(t[i])))+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```

@@

- .whisper[We have sufficient evidence to reject `\\(H_0\\)` in favor of our alternative hypothesis, that there *is a relationship* between class size and test scores].
    -

@@

- .whisper[We are 95% confident that the true relationship is between `\\(-3.22\\)` and `\\(-1.34\\)`.]

@@

- .onfire[If a confidence interval contains the $H_0$ value (i.e. $0$, for our test), then *we fail to reject* $H_0$.]

.footnote[<sup>.red[1]</sup> Technically the `\\(t\\)`-value is `\\(\frac{\hat{\beta_1}-\beta_{1,0}}{se(\hat{\beta_1})}\\)`, but our `\\(H_0\\)` is often that `\\(beta_{1,0}=0\\)`, such that the formula reduces to the one above.]

>>>>

- We want to .onfire[identify] causal relationships between **population** variables
    - Logically first thing to consider
    - **Endogeneity problem**

- We'll use **sample** *statistics* to .onfire[infer] something about population *parameters*
    - In practice, we'll only ever have a finite *sample distribution* of data
    - We *don't* know the *population distribution* of data
    - **Randomness problem**

@@

- .onfire[All modern science is built upon statistical hypothesis testing, so understand it well!]

@@

- Modern philosophy of science is largely based off of hypothesis testing and .onfire[falsifiability], which form the "Scientific Method"<sup>.red[1]</sup>

- For something to be "scientific", it must be .onfire[falsifiable], or at least .onfire[testable]

@@

- .onfire[Test statistic `\\((\delta)\\)`]: measures **how far what we observed in our sample $(\hat{\beta_1})$ is from what we would expect if the null hypothesis were true $(\beta_1=0)$**
    - Calculated from a sampling distribution of the estimator (i.e. $\hat{\beta_1})$
    - In econometrics, we use $t$-distributions which have $n-k-1$ degrees of freedom<sup>.red[1]</sup>

- .onfire[Rejection region]: if the test statistic reaches a .onfire["critical value"] of $\delta$, then we **reject** the null hypothesis

@@

```{r, echo=T}
# save as obs_slope
sample_slope <- school_reg_tidy %>% # this is the regression tidied with broom
  filter(term=="str") %>%
  pull(estimate)

@@

```{r, echo=T, eval=F}
CASchool %>%
  specify(testscr ~ str)
```

@@

```{r}
CASchool %>%
  specify(testscr ~ str) %>%
  hypothesize(null = "independence") %>%
  head(.,3)
```

.footnote[<sup>.red[1]</sup> `type` can be either `point` (for specific point estimates for a single variable, such as a sample mean, `\\((\bar{x})\\)`, or `independence` (for hypotheses about two samples or a relationship between variables). See more [here](https://moderndive.netlify.com/9-hypothesis-testing.html).]

@@

- .onfire[Two-sided alternative] $H_a: \beta_1 \neq 0$, we double the raw $p$-value

@@

.footnote[<sup>1</sup> `\\(k\\)` is the number of `\\(X\\)` variables.]

@@

```{r, fig.retina=3, fig.height=5, fig.align="center"}
ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  stat_function(fun=dnorm, xlim=c(-4,-2), geom="area", fill="blue", alpha=0.5)+
  stat_function(fun=dnorm, xlim=c(2,4), geom="area", fill="blue", alpha=0.5)+
  labs(x = expression(paste("Test Statistic t, (deviation from ", H[0], " if ", H[0], " were True)")),
       y = "Probability")+
  scale_x_continuous(breaks=c(-2,2),
                      label=c(expression(-t[i]), expression(t[i])))+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```

@@

$$\begin{align*}
H_0: \,& \beta_1=0\\
H_a: \, & \beta_a \neq 0\\
\end{align*}$$

- Because the hypothesis test's $p$-value $<$ $\alpha$ (0.05)...

- .onfire[We have sufficient evidence to reject `\\(H_0\\)` in favor of our alternative hypothesis. Our sample suggests that there *is a relationship* between class size and test scores].

@@

- Using the confidence intervals:

- .onfire[We are 95% confident that the true marginal effect of class size on test scores is between `\\(-3.22\\)` and `\\(-1.34\\)`.]

@@

- .whisper[If a confidence interval contains the `\\(H_0\\)` value (i.e. `\\(0\\)`, for our test), then *we fail to reject* `\\(H_0\\)`.]

.footnote[<sup>.red[1]</sup> Since our null hypothesis is that $\beta_{1,0}=0$, the test statistic simplifies to this neat fraction.]

# Common Misconceptions about p-values

- All of the following are **FALSE** interpretations of $p$ (and below are reasons why each is wrong)

1. $p$ **is the probability that the alternative hypothesis is false**
  - We can never *prove* an alternative hypothesis, only tentatively reject a null hypothesis

2. $p$ **is the probability that the null hypothesis is true**
  - We're not *proving* the $H_0$ is false, only saying that it's very unlikely that if $H_0$ were true, we'd obtain a slope as rare as our sample's slope

3. $p$ **is the probability that our observed effects were produced purely by random chance**
  - $p$ is computed under a specific model (think about our null world) that *assumes* $H_0$ is true

4. $p$ **tells us how significant our finding is**
  - $p$ tells us nothing about the *size* or the *real world significance* of any effect deemed "statistically significant"
  - it only tells us that the slope is statistically significantly different from 0 (if $H_0$ is $\beta_1=0)$

# Abusing p-Values I

.pull-left[
.center[
![:scale 60%](https://www.dropbox.com/s/9nv7ewnkokuek3c/smbc1623-1.png?raw=1)
]
]

.pull-right[
.center[
![:scale 60%](https://www.dropbox.com/s/5xdh79z3uwg5n3b/smbc1623-2.png?raw=1)
]
]

.source[Source: [SMBC](http://www.smbc-comics.com/?id=1623)]

# Abusing p-Values II

.center[
![](https://www.dropbox.com/s/70gxtqm7tmag454/asalogo.jpg?raw=1)
]
]

>"The widespread use of 'statistical significance' (generally interpreted as $(p \leq 0.05)$ as a license for making a claim of a scientific finding (or implied truth) leads to considerable distortion of the scientific process."

.source[Wasserstein, Ronald L. and Nicole A. Lazar, (2016), ["The ASA's Statement on p-Values: Context, Process, and Purpose](http://www.tandfonline.com/doi/full/10.1080/00031305.2016.1154108)," *The American Statistician* 30(2): 129-133]

# p-value Clarification

- Again, .shout[p-value is the probability that, assuming the null hypothesis is true, we obtain (by pure random chance) a test statistic at least as extreme as the one we estimated for our sample]

- A low p-value means either (and we can't distinguish which):
    1. $H_0$ is true and a highly improbable event has occurred OR
    2. $H_0$ is false
